---
title: "Models of Neighborhood Change in YOUR CITY NAME"
output: 
  flexdashboard::flex_dashboard:
    theme: bootstrap
    source: embed
    smart: false
runtime: shiny
---



```{r setup, include=FALSE}
knitr::opts_chunk$set( echo = TRUE, message=F, warning=F, fig.width=10, fig.height = 6 )
```


```{r, include=FALSE}

# PACKAGES 


# dashboard layout, widgets, and uploading

library( flexdashboard )
library( shiny )
library( rsconnect )

# data wrangling 

library( dplyr )


# formatting output

library( DT )
library( pander )
library( knitr )
library( stargazer )


# maps

library( ggmap )
library( leaflet )
library( viridis )
library( geojsonio )
library( sp )
library( sf )
library( tmap )
library( pals )
library( rgdal )

####
library(mclust)
library(ggplot2)
library(ggthemes)
library(tidycensus)
library(cartogram)
library(maptools)
library(stringr)
library(fun)

```



```{r, include=FALSE}

crosswalk <- read.csv( "https://raw.githubusercontent.com/DS4PS/cpp-529-master/master/data/cbsatocountycrosswalk.csv",  stringsAsFactors=F, colClasses="character" )

# search for citie names by strings, use the ^ anchor for "begins with" 

grep( "^TOL", crosswalk$msaname, value=TRUE ) 

```



```{r}
these.tol <- crosswalk$msaname == "TOLEDO, OH"
these.fips <- crosswalk$fipscounty[ these.tol ]
these.fips <- na.omit( these.fips )

head( these.fips ) %>% pander()
```


```{r}
state.fips <- substr( these.fips, 1, 2 )
county.fips <- substr( these.fips, 3, 5 )

cbind( these.fips, state.fips, county.fips ) %>% pander()
```

```{r}
tol.pop <-
get_acs( geography = "tract", variables = "B01003_001",
         state = "39", county = county.fips[state.fips=="39"], geometry = TRUE ) %>% 
         select( GEOID, estimate ) %>%
         rename( POP=estimate )
```

```{r}

```




```{r}

tol <- tol.pop

# make sure there are no empty polygons
tol <- tol[ ! st_is_empty( tol ) , ]

# convert sf map object to an sp version
tol.sp <- as_Spatial( tol )

class( tol.sp )

```



```{r}
plot( tol.sp )
```


```{r}
# project map and remove empty tracts
tol.sp <- spTransform( tol.sp, CRS("+init=epsg:3395"))
tol.sp <- tol.sp[ tol.sp$POP != 0 & (! is.na( tol.sp$POP )) , ]

# convert census tract polygons to dorling cartogram

tol.sp$pop.w <- tol.sp$POP / 9000    # standardizes it to max of 1.5
tol_dorling <- cartogram_dorling( x=tol.sp, weight="pop.w", k=0.05 )
plot( tol_dorling )

```
```{r}
#class(tol_dorling)
```

```{r}
# user-defined bounding box to move slocer to subjects 
bb <- st_bbox( c( xmin =  -9401605, xmax = -9260415, 
                  ymax = 5108999, ymin = 5003788 ), 
               crs = st_crs("+init=epsg:3395"))

tm_shape( tol_dorling, bbox=bb ) + 
  tm_polygons( col="POP", n=10, style="quantile", palette="Spectral" ) +
  tm_layout( "Dorling Cartogram", title.position=c("right","top") )

```



```{r}
keep.these <- c("pnhwht12", "pnhblk12", "phisp12", "pntv12", "pfb12", "polang12", 
"phs12", "pcol12", "punemp12", "pflabf12", "pprof12", "pmanuf12", 
"pvet12", "psemp12", "hinc12", "incpc12", "ppov12", "pown12", 
"pvac12", "pmulti12", "mrent12", "mhmval12", "p30old12", "p10yrs12", 
"p18und12", "p60up12", "p75up12", "pmar12", "pwds12", "pfhh12")

#eman

dat1 <- tol_dorling@data

URL <- "https://github.com/DS4PS/cpp-529-master/raw/master/data/ltdb_std_2010_sample.rds"
census.dat <- readRDS(gzcon(url( URL )))

# can merge an sf object and data.frame
dat1 <- merge( dat1, census.dat, by.x="GEOID", by.y="tractid" )


dat2 <- select( dat1, keep.these )
dat3 <- apply( dat2, 2, scale )
head( dat3[,1:6] ) %>% pander()

```


```{r}
set.seed( 1234 )
fit <- Mclust( d3 )
tol_dorling$cluster <- as.factor( fit$classification )
summary( fit )
```


```{r}
URL1 <- "https://github.com/DS4PS/cpp-529-fall-2020/raw/main/LABS/data/rodeo/LTDB-2000.rds"
d1 <- readRDS( gzcon( url( URL1 ) ) )

URL2 <- "https://github.com/DS4PS/cpp-529-fall-2020/raw/main/LABS/data/rodeo/LTDB-2010.rds"
d2 <- readRDS( gzcon( url( URL2 ) ) )

URLmd <- "https://github.com/DS4PS/cpp-529-fall-2020/raw/main/LABS/data/rodeo/LTDB-META-DATA.rds"
md <- readRDS( gzcon( url( URLmd ) ) )

d1 <- select( d1, - year )
d2 <- select( d2, - year )

d <- merge( d1, d2, by="tractid" )
d <- merge( d, md, by="tractid" )
```


```{r}
d <- select( d, tractid, 
             mhmval00, mhmval12, 
             hinc00, 
             hu00, vac00, own00, rent00, h30old00,
             empclf00, clf00, unemp00, prof00,  
             dpov00, npov00,
             ag25up00, hs00, col00, 
             pop00.x, nhwht00, nhblk00, hisp00, asian00,
             cbsa, cbsaname )

 
d <- 
  d %>%
  mutate( # percent white in 2000
          p.white = 100 * nhwht00 / pop00.x,
          # percent black in 2000
          p.black = 100 * nhblk00 / pop00.x,
          # percent hispanic in 2000
          p.hisp = 100 * hisp00 / pop00.x, 
          # percent asian in 2000
          p.asian = 100 * asian00 / pop00.x,
          # percent high school grads by age 25 in 2000 
          p.hs = 100 * (hs00+col00) / ag25up00,
          # percent pop with college degree in 2000
          p.col = 100 * col00 / ag25up00,
          # percent employed in professional fields in 2000
          p.prof = 100 * prof00 / empclf00,
          # percent unemployment  in 2000
          p.unemp = 100 * unemp00 / clf00,
          # percent of housing lots in tract that are vacant in 2000
          p.vacant = 100 * vac00 / hu00,
          # dollar change in median home value 2000 to 2010 
          pov.rate = 100 * npov00 / dpov00 )


# adjust 2000 home values for inflation 
mhv.00 <- d$mhmval00 * 1.28855  
mhv.10 <- d$mhmval12


# change in MHV in dollars
mhv.change <- mhv.10 - mhv.00


# drop low 2000 median home values
# to avoid unrealistic growth rates.
#
# tracts with homes that cost less than
# $1,000 are outliers
mhv.00[ mhv.00 < 1000 ] <- NA


# change in MHV in percent
mhv.growth <- 100 * ( mhv.change / mhv.00 )

#Omit cases with growth rates above 200%.
mhv.growth[ mhv.growth > 200 ] <- NA

d$mhv.00 <- mhv.00
d$mhv.10 <- mhv.10
d$mhv.change <- mhv.change
d$mhv.growth <- mhv.growth

p.vacant<-d$p.asian
p.unemp<-d$p.unemp
p.asian<-d$p.asian
```




```{r}
# STANDARDIZE GEO IDs

# note the current geoid format for the LTDB census data: 
# FIPS-STATE-COUNTY-TRACT:  fips-01-001-020100  

x <- d$tractid 
head( x )
```

```{r}

# remove non-numeric strings 
x <- gsub( "fips", "", x )
x <- gsub( "-", "", x )
head( x )
```

```{r}
# drop leading zeros 
x <- as.numeric( x )

# remember to add the variable back to the census dataset
d$tractid2 <- x 

#???????????which line is correct?????? 
tole <- merge( tol_dorling, d, by.x="GEOID", by.y="tractid", all.x=T ) 
#tol <- merge( tol, d, by.x="GEOID", by.y="tractid", all.x=T )
head(tole@data) %>% pander()
```


```{r}
names(tole) %>% sort()
```


```{r}
# data frame and polygon ID standardization in case a tract was dropped and IDs don't match
row.ids <- sapply( slot( tole, "polygons" ), function(x) slot( x, "ID" ) )
row.names( tole ) <- row.ids

```


```{r}
# project to standard lat-lon coordinate system 
tole <- spTransform( tole, CRS("+proj=longlat +datum=WGS84") )

# write to file 
geojson_write( tole, file="Toledo_dorling.geojson", geometry="polygon" )

```



```{r, include=FALSE}

# DATA STEPS 

# load dorling cartogram from github
# map already contains census data and groups from clustering 

github.url <- "https://github.com/DS4PS/cpp-529-fall-2020/raw/main/data/phx_dorling.geojson"
tol <- geojson_read( x=github.url,  what="sp" )

# reproject the map 
tol2 <- spTransform( tol, CRS("+init=epsg:3395") )

# convert the sp map format to 
# an sf (simple features) format:
# ggmap requires the sf format
tol.sf <- st_as_sf( tol2 )

# separate out the data frame from the map
d <- as.data.frame( tol.sf )



```


Community Demographics  
=====================================  

Inputs {.sidebar}
-------------------------------------

```{r}
these.variables <- c("pnhwht12", "pnhblk12", "phisp12", "pntv12", "pfb12", "polang12", 
"phs12", "pcol12", "punemp12", "pflabf12", "pprof12", "pmanuf12", 
"pvet12", "psemp12", "hinc12", "incpc12", "ppov12", "pown12", 
"pvac12", "pmulti12", "mrent12", "mhmval12", "p30old12", "p10yrs12", 
"p18und12", "p60up12", "p75up12", "pmar12", "pwds12", "pfhh12")


# replace these with descriptive labels 
# from the data dictionary 
temp.names <- paste0( "Variable ", these.variables )

radioButtons( inputId="demographics", 
              label = h3("Census Variables"),
              # choices = these.variables, 
              choiceNames=temp.names,
              choiceValues=these.variables,
              selected="pnhwht12")



# Adding interpretable variable names
# from the data dictionary:
# add a name attribute for each variable
# 
# value <- c(1,2,3)
# dd.name <- c("one","two","three")
# 
# x <- dd.name
# names(x) <- value
#
# dd names and values linked
# names( x[2] )
#
# can now get the label using the value
# using the name attributes 
# x[ "two" ]
#
# to add labels to the maps
# use the radio button value 
# to get the data dictionary label: 
#
# x[ input$demographics ]
```



Row {.tabset}
-------------------------------------


### Choropleth Map


```{r}

renderPlot({

# split the selected variable into deciles 
  
get_data <- 
  reactive({
             tol.sf <- 
             tol.sf %>% 
             mutate( q = ntile( get(input$demographics), 10 ) )  
          })

  
ggplot( get_data() ) +
    geom_sf( aes( fill = q ), color=NA ) +
    coord_sf( datum=NA ) +
    labs( title = paste0( "Choropleth of Select Demographics: ", toupper(input$demographics) ),
          caption = "Source: Harmonized Census Files",
          fill = "Population Deciles" ) +
    scale_fill_gradientn( colours=rev(ocean.balance(10)), guide = "colourbar" ) + 
    xlim( xmin = -12519146, xmax = -12421368 ) + 
    ylim( ymin = 3899074, ymax = 3965924 )

})

```


### Variable Distribution 

```{r}
renderPlot({

# extract vector x from the data frame 
# x <-  d[ "pnhwht12" ] %>% unlist()

  
get_variable_x <- reactive({ d[ input$demographics ] })

x <- get_variable_x() %>% unlist()

cut.points <- quantile( x, seq( 0, 1, 0.1 ) )

hist( x, breaks=50, 
      col="gray", border="white", yaxt="n",
      main=paste0( "Histogram of variable ", toupper( input$demographics ) ),
      xlab="red lines represent decile cut points" )

abline( v=cut.points, col="darkred", lty=3, lwd=2 )


})
```




Neighborhoods  
===================================== 

### Clusters  

```{r}

# define the bounding box corners 
bb <- st_bbox( c( xmin = -12519146, xmax = -12421368, 
                  ymax = 3965924, ymin = 3899074 ), 
               crs = st_crs("+init=epsg:3395"))


# ADD YOUR CUSTOM LABELS TO THE CLUSTERS 

tol2$cluster[ tol2$cluster == "1" ] <- "Baby Boomers"
tol2$cluster[ tol2$cluster == "2" ] <- "Hipsters"



renderTmap({

tmap_mode("view")
tm_basemap( "CartoDB.Positron" )
tm_shape( phx2, bbox=bb ) + 
  tm_polygons( col="cluster", palette="Accent",
               title="Community Types" ) 

})
```




NH Change 2000-2010  
===================================== 



Inputs {.sidebar}
-------------------------------------

```{r}

button.labels <- c("Median Home Value 2000","Median Home Value 2010","Value Change 2000-2010","Growth in Home Value")
button.values <- c("mhv.2000","mhv.2010","mhv.change","mhv.growth")
  
radioButtons( inputId="home.value", 
              label = h3("Home Values"),
              # choices = these.variables, 
              choiceNames=button.labels,
              choiceValues=button.values,
              selected="mhv.2000")

```




Row {.tabset}
-------------------------------------



### Median Home Values


```{r}

renderPlot({

# split the selected variable into deciles 
  
get_data <- 
  reactive({
             tol.sf <- 
             tol.sf %>% 
             mutate( q = ntile( get(input$home.value), 10 ) )  
          })

  
ggplot( get_data() ) +
    geom_sf( aes( fill = q ), color=NA ) +
    coord_sf( datum=NA ) +
    labs( title = paste0( "Spatial Distribution of Home Values: ", toupper(input$demographics) ),
          caption = "Source: Harmonized Census Files",
          fill = "Home Value Deciles" ) +
    scale_fill_gradientn( colours=rev(ocean.balance(10)), guide = "colourbar" ) + 
    xlim( xmin = -12519146, xmax = -12421368 ) + 
    ylim( ymin = 3899074, ymax = 3965924 )

})

```


### Variable Distribution 

```{r}
renderPlot({

# extract vector x from the data frame 
# x <-  d[ "pnhwht12" ] %>% unlist()

  
get_variable_x <- reactive({ d[ input$home.value ] })

x <- get_variable_x() %>% unlist() %>% as.numeric()

cut.points <- quantile( x, seq( 0, 1, 0.1 ) )

hist( x, breaks=50, 
      col="gray", border="white", yaxt="n",
      main=paste0( "Histogram of ", toupper( input$home.value ) ),
      xlab="red lines represent decile cut points" )

abline( v=cut.points, col="darkred", lty=3, lwd=2 )


})
```  







Drivers of Change   
===================================== 



Inputs {.sidebar}
-------------------------------------

```{r}

button.labels <- c("Median Home Value 2000","Median Home Value 2010","Value Change 2000-2010","Growth in Home Value")
button.values <- c("mhv.2000","mhv.2010","mhv.change","mhv.growth")
  
radioButtons( inputId="dv", 
              label = h3("Select Your Dependent Variable"),
              choiceNames=button.labels,
              choiceValues=button.values,
              selected="mhv.change")


covariates <- c("pnhwht12", "pnhblk12", "phisp12", "pntv12", "pfb12", "polang12", 
"phs12", "pcol12", "punemp12", "pflabf12", "pprof12", "pmanuf12", 
"pvet12", "psemp12", "hinc12", "incpc12", "ppov12", "pown12", 
"pvac12", "pmulti12", "mrent12", "mhmval12", "p30old12", "p10yrs12", 
"p18und12", "p60up12", "p75up12", "pmar12", "pwds12", "pfhh12")

# covariate.labels <- c( ... )

checkboxGroupInput( inputId="covariates", 
              label = h3("Select Variables for Your Model"),
              choices = covariates,
              # choiceNames=covariate.labels,
              # choiceValues=covariates,
              selected=c("pnhwht12","pprof12","pvac12") )

```




Row {.tabset}
-------------------------------------



### Predicting Change 

```{r, results="asis"}

# RUNNING A REGRESSION WITH USER INPUTS
#
# create a formula object
# by constructing the formula from user selections 
# as a string then casting as a formula object 
# x.s <- c("x1","x2","x3" )
# formula.text <- paste( "y", " ~ ", paste0( x.s, collapse=" + ") )
# formula.object <- as.formula( formula.text )
#
# lm( formula.object, data=d )
#
# make sure all variables are in d


  
# check boxes return vectors 
get_covariates <-  reactive({ input$covariates })


renderUI({
  
covariates <- get_covariates() 

formula.text <- paste0( "mhmval12", " ~ ", paste( covariates, collapse=" + " ) )
fo <- as.formula( formula.text )

m <- lm( fo, data=d )


# HTML( "<div style='width: 60%; margin: 0px auto;'>" )
HTML( 
  
  c("<br><br><br>",
  "<div type='regression' style='width: 60%; margin: 0px auto;'>",
  stargazer( m, type="html", omit.stat=c("rsq","f") ),
  "</div>",
  "<br><br><br>"
  )
  
  )


})

# HTML( reg.table )
```



### Correlation Plots 

```{r}
pairs( iris )
```





<style>

.chart-shim { overflow: auto; }
 
table{
   border-spacing:1px;
   margin-top:30px;
   margin-bottom:30px;
   margin-left: auto;
   margin-right: auto;
   align:center} 

td{ padding: 6px 10px 6px 10px } 

th{ text-align: left; } 

</style>
